Modal Abstractions for Virtualizing Memory Addresses

### 2025-07-16 background.tex
experiment w/ paragraph giving more detail on mapping a page, in bg section

>>>>

\paragraph{\add{Mapping a New Page}}
\add{Mapping a new page into a virtual address space is a key piece fo functionality, and also a key challenge for verification.
This operation takes a chunk of physical memory to install, and a (page-aligned) virtual address at which to make it
visible to the process.
Starting from the page table root (\lstinline|cr3| in Figure \ref{fig:pagetables}), it must traverse the page tables until it locates
the correct location of the L1 entry which should point to the physical page (installing new L3, L2, and L1 tables as needed),
and update the L1 entry to point to the physical page. This is complicated by the fact that the kernel accesses the page
tables through \emph{virtual} addresses, but the page table root and the addresses in the L4--L1 entries are all \emph{physical}
addresses. So the mapping process must repeatedly perform \emph{backwards} mappings (not supported by hardware) from
physical addresses of each level's entry to a virtual address that is then mapped back to the corresponding physical address!
And for modularity, the translation process (and possible allocation of missing tables) is typically implemented in a function
used for every level of translation (i.e., a level-polymorphic function).
These compounding subtleties are one reason \textsc{CertiKOS} and \textsc{seL4} \emph{trust} their page table
manipulations; both the reasoning to prove correctness, and the kernel invariants that ensure correctness of typical implementations,
are remarkably subtle.
}
\looseness=-1
